---
title: "Realistic scenario (Ge et al. 2011)"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
source("R/utils/sim_data_utils.R")
source("R/utils/sim_reporting_utils.R")
source("R/utils/sim_variance_estimators.R")
```

## Realistic scenario

We consider the data generating mechanism from Ge et al. (2011) as a realistic scenario. As described in the paper:

>The simulated trials were constructed to approximately
match the conditions of the real trial described
in Section 2. 
>
>Each simulated trial had 350 patients with 
2:1 stratified randomization.
>
>A continuous standard normal covariate was
generated along with an independent binary
covariate with 0/1 proportions of 0.50. 
>
>The continuous covariate was categorized by dividing
it at its population median to yield
2 x 2 = 4 strata levels for the design, or it was
divided at its population quartiles to yield
4 x 2 = 8 strata levels. 
>
>The number of patients
in each stratum varied across the simulated trials,
but an approximate 2:1 randomization ratio
was implemented within each stratum (with
rounding to whole numbers when required).
>
>The binary response variable was simulated
based on a logistic regression model computed
using the covariate values in each simulated
data set. 
The base case has coefficients b0 = -1.8, b1 = 1.6,
b2 = -0.6, and b3 = 0.0 corresponding to the
intercept (b0), treatment (b1), continuous covariate
(b2), and binary covariate (b3).


$$
  logit[P(Y_i = 1 | A, X)] = b_0 + b_1A + b_2X_1 + b_3X_2
$$


In this case we replicate scenario 1 (with $b_0 = -1.8$) from the paper with the 4 strata level design. We discretize the continuous covariate ('x1' in the code) at the population median (i.e., zero) to produce stratification factor 'z1'. Treatment assignment is then stratified by 'z1' and the binary covariate 'x2'. 
The outcome model in this scenario corresponds to

$$
  logit[P(Y_i = 1 | A, X)] = -1.8 + 1.6A + -0.6X_1 + 0X_2
$$

Our working model is $logit[P(Y_i = 1 | A, X)] = b_0 + b_1A + b_2Z_1 + b_3X_2$ and thus our results correspond to method 'LR-Discrete-2' in Ge et al (2011). 


```{r}
# Set up scenario

scenarios <- list(
  list(scenario = 1, 
       N = 350,
       
       # continuous (x1) and binary (x2) covariates
       covars = tibble("name"=c("x1", "x2"),
                       "dist"=c("rnorm", "rbinom"),
                       "arg1"=c(0, 1),
                       "arg2"=c(1, 0.5)),
       
       # define strata variable z1 by discretizing x1 at population median
       make_strata = tibble("name" = c("z1"), 
                            "covar" = c("x1"), 
                            "breaks" = c(0)),
       
       # stratified randomization according to x2 and z1 (2x2=4 strata)
       stratified_randomization = T,
       strata = c("x2", "z1"),
       # 2:1 assignment
       trt_assignment = 2/3,
       
       # outcome model uses discretized covariate z1
       response_func = \(df) -1.8 + 1.6*df$trt + -0.6*df$x1 + 0*df$x2,
       working_model = "y ~ trt + z1 + x2"
  )
)

```


The true marginal probabilities that Y=1 are calculated below per stratum ($Z_1$) and treatment arm using $N=10^6$:

```{r}
# True marginal probabilities that Y=1

df <- data.frame(sim_covars(list(), scenarios[[1]], N = 1e6))

# for patients in stratum Z == 0
marginal_00 <- mean(plogis(scenarios[[1]]$response_func(df |>
                               transform(trt = 0) |>
                                 filter(z1 == 0))))
marginal_00

marginal_10 <- mean(plogis(scenarios[[1]]$response_func(df |>
                               transform(trt = 1) |>
                                 filter(z1 == 0))))
marginal_10

# marginal log-odds ratio in stratum Z == 0
log((marginal_10/(1-marginal_10)) / (marginal_00/(1-marginal_00)))


# for patients in stratum Z == 1
marginal_01 <- mean(plogis(scenarios[[1]]$response_func(df |>
                               transform(trt = 0) |>
                                 filter(z1 == 1))))
marginal_01

marginal_11 <- mean(plogis(scenarios[[1]]$response_func(df |>
                               transform(trt = 1) |>
                                 filter(z1 == 1))))
marginal_11

# marginal log-odds ratio in stratum Z == 1
log((marginal_11/(1-marginal_11)) / (marginal_01/(1-marginal_01)))

```



### Unconditional coverage


```{r}

## Unconditional coverage 
## PATE, CPATE

## Simulate 10,000 runs
N_runs <- 1e4
sims <- Q(run_sim_wrapper, 
          scenario=rep(scenarios, each=N_runs), 
          n_jobs = 200)
sims <- do.call(rbind, sims)


var_methods = list("ye_beeca_var",
                   "ye_beeca_strata_var",
                   "ge_var")

# PATE 
post_sims_pate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_pate))
results_pate <- purrr::map_df(post_sims_pate, sum_sims)
results_pate$estimand <- "PATE"


# CPATE
post_sims_cpate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_cpate))
results_cpate <- purrr::map_df(post_sims_cpate, sum_sims)
results_cpate$estimand <- "CPATE"

```

Results: Unconditional coverage for PATE estimand

```{r}
dplyr::arrange(results_pate, scenario)
```

Results: Unconditional coverage for CPATE estimand

```{r}
dplyr::arrange(results_cpate, scenario)
```


### Coverage conditional on X


```{r}

## Conditional coverage (conditional only on X)
## PATE, CPATE

#===============================================================================
## simulate a base dataset that remains fixed
## we fix X, but simulate TRT and y
## (achieved by dropping trt and y from base_data so they are re-simulated each time)

all_res_pate <- list()
all_res_cpate <- list()
i <- 1

for (rseed in c(274)) {
  
  set.seed(rseed)
  
  base_data <- sim_dat(scenarios[[1]])
  base_data <- base_data %>% select(-trt, -y)
  
  #=============================================================================
  
  ## Simulate 10,000 runs
  N_runs <- 1e4
  sims <- Q(run_sim_wrapper, 
            scenario=rep(scenarios, each=N_runs), 
            const = list(base_data = base_data), 
            n_jobs = 200)
  sims <- do.call(rbind, sims)
  
  var_methods = list("ye_beeca_var",
                     "ye_beeca_strata_var",
                     "ge_var")
  
  # PATE with conditional coverage (conditional on X)
  post_sims_pate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_pate))
  results_pate <- purrr::map_df(post_sims_pate, sum_sims)
  results_pate$seed <- rseed
  results_pate$estimand <- "PATE"
  
  
  # CPATE with conditional coverage (conditional on X)
  post_sims_cpate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_cpate))
  results_cpate <- purrr::map_df(post_sims_cpate, sum_sims)
  results_cpate$seed <- rseed
  results_cpate$estimand <- "CPATE"
  
  
  all_res_pate[[i]] <- results_pate
  all_res_cpate[[i]] <- results_cpate
  
  i <- i + 1
  
}

multi_res_pate <- do.call(rbind, all_res_pate)
multi_res_cpate <- do.call(rbind, all_res_cpate)


```

Results: Conditional coverage on X, for PATE estimand

```{r}
multi_res_pate
```

Results: Conditional coverage on X, for CPATE estimand

```{r}
multi_res_cpate
```


### Coverage conditional on A, X


```{r}

## Conditional coverage (conditional on X and TRT)
## PATE, CPATE

#===============================================================================
## simulate a base dataset that remains fixed
## we fix X and TRT, but simulate y
## (achieved by dropping y from base_data so it is re-simulated each time)

all_res_pate <- list()
all_res_cpate <- list()
i <- 1

for (rseed in c(274)) {
  
  set.seed(rseed)
  
  base_data <- sim_dat(scenarios[[1]])
  base_data <- base_data %>% select(-y)
  
  #=============================================================================
  
  ## Simulate 10,000 runs
  N_runs <- 1e4
  sims <- Q(run_sim_wrapper, 
            scenario=rep(scenarios, each=N_runs), 
            const = list(base_data = base_data), 
            n_jobs = 200)
  sims <- do.call(rbind, sims)
  
  
  var_methods = list("ye_beeca_var",
                     "ye_beeca_strata_var",
                     "ge_var")
  
  # PATE with conditional coverage (conditional on X, TRT)
  post_sims_pate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_pate))
  results_pate <- purrr::map_df(post_sims_pate, sum_sims)
  results_pate$seed <- rseed
  results_pate$estimand <- "PATE"
  
  
  # CPATE with conditional coverage (conditional on X, TRT)
  post_sims_cpate <- purrr::map(var_methods, post_sims, sims = sims %>% rename(true_diff = true_diff_cpate))
  results_cpate <- purrr::map_df(post_sims_cpate, sum_sims)
  results_cpate$seed <- rseed
  results_cpate$estimand <- "CPATE"
  
  
  all_res_pate[[i]] <- results_pate
  all_res_cpate[[i]] <- results_cpate
  
  i <- i + 1
  
}

multi_res_pate <- do.call(rbind, all_res_pate)

multi_res_cpate <- do.call(rbind, all_res_cpate)

```


Results: Conditional coverage on A, X, for PATE estimand

```{r}
multi_res_pate
```


Results: Conditional coverage on A, X, for CPATE estimand

```{r}
multi_res_cpate
```





